---
title: "Assignment 2 Regression Modeling Helper Part 1 v1"
author: "Lynd Bacon, lynd.bacon@hsc.utah.edu"
output:
  html_notebook:
    toc: yes
  pdf_document:
    toc: no
  html_document:
    toc: yes
---

Creative Commons CC by 4.0 Lynd Bacon & Associates, Ltd.

### Introduction

What follows is intended to give you a kick start on the regression modeling part of this assignment.  Before you do any modeling you'll need to:

* Input the assignment data
* Do any required transformations of the data, create new variables you need.  For example,you need to crate a binary (1/0 variable) from the responses to q12.

Let's do a trial run here.

### Get the data, create what you need

The survey data are in the file appHappyData-2018.RData.  This is an R system file that has two data frames in it.  One data frame has numerical data, and the other has character data.  We want to work with the former.

Assuming that this data file is in the current working directory:

```{r}
appHappyData=load('appHappyData-2018.RData')
ls()
```
You should see the two apphappy data frames listed.

names(apphappy4.num.frame)
Let's give the data we'll be working with a shorter name, and pick out some variables to use:

```{r}
appHappyNum=apphappy.4.num.frame[,c('q12','q48','q49','q50r1')]
```


Let's make the binary dependent variable we need, now.  Here's one way to do it.
Note that we've inserted it into the data.frame we're working with.  The use of table() provides a quick check on the recoding: 

```{r}
appHappyNum['onlyFreeApps']=as.numeric(appHappyNum$q12>=6)
table(appHappyNum$q12, appHappyNum$onlyFreeApps)
```

### Now for a "starter" logistic regression model

Let's predict _onlyFreeApps_ using q48 and q49.

```{r}
logModel1=glm(onlyFreeApps~q48+q49,data=appHappyNum,family=binomial(link=logit))
summary(logModel1)
```
### Some things to note, to think about

Note how a model formula is specified in R:  `onlyFreeApps~q48+q49`.

For a probit type model, change `link=logit` to `link=probit`.

What is "AIC" in the agove output?

