---
title: "IS6481-Fall2018-Assignment-3: Cluster Stat Comparisons"
author: "Lynd Bacon, lynd.bacon@hsc.utah.edu"
output:
  html_notebook:
    toc: yes
  pdf_document:
    toc: no
  html_document:
    toc: yes
---
Creative Commons CC by 4.0 Lynd Bacon & Associates, Ltd.

### Get the Example Data

The data are in the files `Q24rm.rds` and `km3.rds`.  The latter is a three cluster kmeans solution obtained by clustering the Q24 items in `Q24rm.rds`.

You may need to set your working directory. The following is commented out.

```{r setup}
#knitr::opts_knit$set(root.dir = '/media/sdb2/UT2018/IS6481/work/IS6481-Fall-2018/assignments/assignment-3')
```

```{r}
getwd()
Q24nm=readRDS('Q24nm.rds')
km3=readRDS('km3.rds')
```

Here's what's in these two objects:

```{r}
names(Q24nm)
names(km3)
```

### Comparing Cluster Means

Let's compare the means of the three clusters on the first Q24 item, `Q24r1`.  We'll use the R function `oneway.test`.  See help(oneway.test) in R.

First we'll add the kmeans cluster assignments to the `Q24nm` data.frame to Make Life Easier:

```{r}
Q24nm['cluster']=km3$cluster
```
The above will "work" (not generate an error) if nrow(Q24nm)=length(km3$cluster)

```{r}
oneway.test(q24r1~cluster,data=Q24nm)
```
Another way to do this would be:

```{r}
anova(lm(q24r1~as.factor(cluster),data=Q24nm))
```

### Comparing Category Response Frequencies Across Clusters

The data on discrete variables can't usually be averaged. To compare clusters on a discrete variable, we can "cross-tabulate" cluster membership by the discrete variable.  For example:

```{r}
xtabs(~cluster+q12,data=Q24nm)
```

These counts don't tell us whether the clusters are different across the q12 response categories.  We can do a test to see if they differ:

```{r}
chisq.test(xtabs(~cluster+q12,data=Q24nm))
```
This is a test of whether the columns are "independent" of the rows, or vice versa.



